---
title: "Movies personal project"
author: "Ian"
date: "2022-11-02"
output: html_document
---

```{r setup, include=FALSE}
library("knitr")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set("Documents/Masters/Kaggle practice/")

library(tidyverse)
library(kableExtra)

```

## Looking at IMDb Movies and Genres dataset 

Being a movie junkie, I have a personal interest in exploring this dataset. Some possible points of interests can be

- Big directors and their genres 
- Have Ratings changed over time 
- Is there an interaction between main and side genre that gives higher Ratings 
- Censorship over time 
- Relationships between actors and directors

Potential modelling

- Can we predict ratings given variables
  - Need to first check for relationships 


```{r read dataset, echo=FALSE, results='hide'}
movies <- read.csv("IMDb_All_Genres_etf_clean1.csv")

kable(head(movies)) %>% kable_styling()
kable(str(movies)) %>% kable_styling()
kable(summary(movies)) %>% kable_styling()
```

Let us do a cursory check on the data 

```{r cleaning data}
kable(sapply(movies, typeof)) %>% kable_styling()
```
We have a few columns: `r names(movies)`, and we have a total of `r nrow(movies)` rows of data (i.e. 5562 movies)

We can see that something looks off - for example, total_gross should not be a character, lets take a look at a sample: `r head(movies$Total_Gross)` 
- we can already see that it is probably made into a string due to presence of "Gross Unknown": `r head(unique(movies$Total_Gross), 10)`
- hence we need to clean this up 

We may also want to change some columns to factor, such as censor and genres 

```{r cleaning, results='hide'}
# sum(movies$Total_Gross == "Gross Unkown")

## replace "Gross Unkown" with NA
movies[movies$Total_Gross == "Gross Unkown",]$Total_Gross = NA

## remove the "$" and "M" prefix and suffix from the non-problematic entries
library(stringr)
movies[!is.na(movies$Total_Gross),]$Total_Gross <- 
  str_sub(movies[!is.na(movies$Total_Gross),]$Total_Gross, 2, -2)

## make numeric 
movies$Total_Gross <-as.numeric(movies$Total_Gross)

## set Censorship, Main genre to factor
movies$Censor <- as.factor(movies$Censor)
movies$main_genre <- as.factor(movies$main_genre)
```

## Univariate analysis

To explore the data, we can look at some columns of interest one by one 

```{r explore column 1}
movies %>% ggplot(aes(x = Year)) + geom_histogram(col = "red") +
  theme_classic() + ## for a nicer background
  ggtitle(label = "Movie Years", subtitle = "When were the movies made") +
  theme(plot.title = element_text(hjust = .5),
        plot.subtitle = element_text(hjust = .5))
```
The years for the movies in this dataset are left skewed -- there are fewer movies made in the past than closer to the present. 
- The range of years are `r range(movies$Year)`, representing more than 100 years of movies 


```{r explore column 2}
top10directors <- movies %>% select(Director) %>% count(Director) %>% arrange(desc(n)) %>%
  top_n(n = 10)

## reorder the x in terms of -n in order to arrange bars in descending order 
ggplot(top10directors, aes(x = reorder(Director, -n), y = n, fill = n)) + 
  theme_bw() +
  geom_bar(stat = "identity", color = "black")+
  labs(title = "Directors with most movies", 
       x = "Directors",
       y = "Count")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = .5))  ## more visible tick markets for x axis 
```

```{r col 3, warning=FALSE}
## vectors of all actors in the dataset 
actorlist <- unlist(strsplit(movies$Actors, ",")) ## split according to commas
movies$Actors <- strsplit(movies$Actors, ",")

movies$Actors <- str_squish(movies$Actors) ## remove whitespace from front 

## get 20 biggest actors in terms of number of movies 
top20actorstable <- sort(table(Actors = actorlist), decreasing = T)[1:20]

## one hot encoding 
top20actorsnames <- names(top20actorstable)

movies$bigactor <- NA

for (i in nrow(movies)){
  listactors <- unlist(movies[i,]$Actors)
  if (any(is.element(listactors, top20actorsnames))){
    movies[i,]$bigactor <- 1 
  }
  else {movies[i,]$bigactor <- 0}
}

```

`r kable(top20actorstable, format = "html") %>% kable_styling()`


```{r col 4}
movies %>% ggplot(aes(Rating)) + geom_histogram(col = "red") +
  theme_bw() +
  geom_vline(xintercept = mean(movies$Rating), linetype = "dashed")
```
```{r}
movies %>% ggplot(aes(Runtime.Mins.)) + geom_histogram(col = "red") +
  theme_bw() +
  geom_vline(xintercept = mean(movies$Runtime.Mins.), linetype = "dashed")

```

Very right skewed movie durations, with most centering around low 100 minutes. 

- Range of values: there are obviously outliers, `r range(movies$Runtime.Mins.)`, with the highest at 321, `r 321/60` hours of movies, which is very long! 

- The movie(s) is(are) `r movies[movies$Runtime.Mins.>300,]$Movie_Title`

```{r investigate gross and also genre as a byproduct}
sum(is.na(movies$Total_Gross))

missinggross_movies <- movies[is.na(movies$Total_Gross),]

## Investigate?

# in terms of Year
movies %>% ggplot(aes(x = Year)) + geom_histogram(aes(fill = is.na(Total_Gross)), col = "black") +
  scale_fill_manual(values = c("#00BFC4","#F8766D" ))+ ## flip the colours 
  guides(fill = guide_legend(title = "missing data?"))

moviesgenrecount <- movies %>% group_by(main_genre) %>% count()

## table of missing vs not missing genre (for plotting barplots)
table_plot <- movies %>% group_by(main_genre) %>% summarise(no_missing = sum(is.na(Total_Gross))) %>% left_join(moviesgenrecount) %>% mutate(no_not_missing = n - no_missing) %>% select(main_genre, no_missing, no_not_missing)

## make long table to be able to plot on ggplot barplot
longtable <- table_plot %>%pivot_longer(cols = c("no_missing", "no_not_missing"), names_to = "Missing or not",
                           values_to = "Count")

longtable %>% ggplot(aes(x = reorder(main_genre, -Count), y = Count, fill = `Missing or not`)) + geom_bar(stat = "identity", position = "stack") +labs(title = "Genres with most missing gross data", 
       x = "Genres",
       y = "Count")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = .5)) +
  scale_fill_discrete(labels = c("Missing", "Not Missing")) ## change legend labels in scale_fill_discrete
  
```

From the histogram, there is a skew towards modern day movies that are missing gross totals. 

In terms of directors, `r sum(is.element(missinggross_movies$Director, top10directors))` of the missing ones come from the top 10 directors. 

Missing data for gross come somewhat proportionate from genres as well. 

```{r censorship data}
kable(table(movies$Censor))
```

